Climate-RWI relations

### 2018-01-04 analysis/rwi/climate_rwi_relate.Rmd
add spei correlations

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("tidyverse")
library("stringr")
library("treeclim")
```

@@

```{r}
eobsN <- eobs %>% filter(year < 2017) %>% filter(loc == 'N')

prec_N <- eobsN %>% dplyr::select(year, month, prec) 
tmean_N <- eobsN %>% dplyr::select(year, month, tmean) 
tmin_N <- eobsN %>% dplyr::select(year, month, tmin) 
tmax_N <- eobsN %>% dplyr::select(year, month, tmax)

eobsS <- eobs %>% filter(year < 2017) %>% filter(loc == 'S')

prec_S <- eobsS %>% dplyr::select(year, month, prec) 
tmean_S <- eobsS %>% dplyr::select(year, month, tmean) 
tmin_S <- eobsS %>% dplyr::select(year, month, tmin) 
tmax_S <- eobsS %>% dplyr::select(year, month, tmax) 
```

# Precipitation
from previous June to

```{r}
rprec_SJ <- dcc(chrono = cro_sj, climate = prec_N, 
         method = "correlation",  boot = "std",
        .range(-3:9) + .sum(-9:8))

@@

rprecs <- c('rprec_SJ', 'rprec_caH', 'rprec_caL')
rprec_all <- c()

for (i in rprecs){

  # Get object dcc
  aux <- get(i)

  # Get coeffs 
  aux_coef <- aux$coef

  # remove rownames
  row.names(aux_coef) <- NULL

  # Change name for aggregate variables 
  aux_coef <- aux_coef %>% 
    mutate(month = case_when(
      month == "sep...AUG" ~ "Hydrol", 
      TRUE ~ as.character(.$month)),
      sig = case_when(
      significant == "TRUE" ~ "*", 
      significant == "FALSE" ~ ""))

  # Add loc variable
  name_site <- str_sub(i, start=7)
  aux_coef$site <- name_site

  # rbind
  rprec_all <- rbind(rprec_all, aux_coef)
  } 
```

```{r}

rprec_all

rprec_all %>% ggplot(aes(x=id, y=coef, group = site)) +
  geom_bar(aes(fill=site),
           stat="identity", position = "dodge") +
  scale_x_continuous(breaks = rprec_all$id, labels = rprec_all$month) +
  theme_bw() + 
  theme(axis.title.x = element_blank()) +
  geom_text(aes(label = sig), 
            position = position_dodge(0.9), 
            vjust = 0.1, 
            size = 10) +
  ylab('Correlation')

```{r}

eobs_norht_no_prec <- eobs_norht %>% dplyr::select(-prec)

mc <- dcc(chrono = cro_sj, 
          climate = eobs_norht, 
          var_names = c("prec", "tmax", "tmean", "tmin"), 
          method = "correlation", boot = "stationary")

sea_sj <- seascorr(chrono = cro_sj, 
             climate = eobs_norht, 
             var_names = c("prec", "tmax", "tmean", "tmin"))

plot(sea_sj)

dc_sj

croN <- cro %>% filter(site == 'sj') %>% dplyr::select(-site) %>% 
  filter(year > 1949)
rownames(croN) <- croN$year

cl <- eobs %>% 
  filter(loc == 'N') %>% 
  filter(variable %in% c('rr', 'tg')) %>% 
  dplyr::select(value, variable, year, month) %>% 
  spread(variable, value)

cl <- cl %>% rename(prec = rr, temp = tg) %>% 
  filter(year < 2017)

test <- dcc(chrono = croN, climate = cl)

croN 
# http://czang.org/treeclim.html

>>>>

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library("tidyverse")
library("stringr")
library("treeclim")
# devtools::install_github("thomasp85/patchwork")
library("patchwork")
```

@@

```{r}
eobsN <- eobs %>% filter(year < 2017) %>% filter(loc == 'N') 
prec_N <- eobsN %>% dplyr::select(year, month, prec) 
tmean_N <- eobsN %>% dplyr::select(year, month, tmean) 
tmin_N <- eobsN %>% dplyr::select(year, month, tmin) 
tmax_N <- eobsN %>% dplyr::select(year, month, tmax)

eobsS <- eobs %>% filter(year < 2017) %>% filter(loc == 'S') 
prec_S <- eobsS %>% dplyr::select(year, month, prec) 
tmean_S <- eobsS %>% dplyr::select(year, month, tmean) 
tmin_S <- eobsS %>% dplyr::select(year, month, tmin) 
tmax_S <- eobsS %>% dplyr::select(year, month, tmax) 
```

```{r}
eobs <- read.csv(file=paste0(di, "data/eobs/eobs_formatted.csv"), header=TRUE, sep=',')

# sanjuan: 37.133, -3.382 = spei_-3.25_37.25.csv 
# canar: 36.957, -3.4284 = spei_-3.25_36.75.csv

spei_sj <- read.csv(file=paste0(di, "data_raw/spei/single_location/SPEI_37.25_-3.25.csv"), header=TRUE)
spei_ca <- read.csv(file=paste0(di, "data_raw/spei/single_location/SPEI_36.75_-3.25.csv"), header=TRUE)

# Prepare data 
spei_N <- spei_sj %>% 
  mutate(date = as.Date(paste0('01', DATA), format =  '%d%b%Y'),
         year = lubridate::year(date),
         month = lubridate::month(date)) %>% 
  dplyr::select(year, month, spei6 = SPEI_6) %>% 
  filter(year > 1950) %>% 
  filter(year < 2017)

spei_S <- spei_ca %>% 
  mutate(date = as.Date(paste0('01', DATA), format =  '%d%b%Y'),
         year = lubridate::year(date),
         month = lubridate::month(date)) %>% 
  dplyr::select(year, month, spei6 = SPEI_6) %>% 
  filter(year > 1950) %>% 
  filter(year < 2017)
```

## Some useful functions

```{r}
# for join dccs 
join_dccs <- function(dfs){ 
  out <- c() 
  for (i in dfs){
    # Get object dcc
    aux <- get(i)
    # Get coeffs 
    aux_coef <- aux$coef
    # remove rownames
    row.names(aux_coef) <- NULL

    # # Add loc variable
    # name_site <- sub(".*_", "", i)
    # aux_coef$site <- name_site

    # Add variable sig (asterisk) and site 
    aux_coef <- aux_coef %>% 
      mutate(site = sub(".*_", "", i), 
            sig = case_when(
              significant == "TRUE" ~ "*", 
              significant == "FALSE" ~ ""))

    # rbind
    out <- rbind(out, aux_coef)

  } 
  return(out)
}

## PLOTs functions 
theme_cor <- function(){
  theme_bw() %+replace%
    theme(axis.title.x = element_blank(),
          panel.grid = element_blank(),
          axis.text.x = element_text(angle = 90))
}
# http://joeystanley.com/blog/custom-themes-in-ggplot2

custom_gg <- list(theme_cor(), 
                  geom_hline(yintercept = 0, colour="gray"),
                  ylab('Correlation'), 
                  geom_bar(aes(fill=site), stat="identity", position = "dodge"),
                  geom_text(aes(label = sig, 
                                x = id, 
                                y = (coef + ifelse(coef >= 0, 0.02, -0.02))), 
                            position = position_dodge(.9), size = 10),
                  scale_fill_manual(values = c("#B2DF8A","#1F78B4","#A6CEE3")))
                  # scale_fill_brewer(palette = 'Paired', type='qual'))

# "#A6CEE3" azul clarito
# "#B2DF8A" verde

# Precipitation

* Correlación entre Marzo del mes anterior y hasta Sep del mes en curso
* Correlación con datos de Pp acumulada

```{r}
set.seed(3333)
rprec_SJ <- dcc(chrono = cro_sj, climate = prec_N, 
         method = "correlation",  boot = "std",
        .range(-3:9) + .sum(-9:8))

@@

# Joins dccs 
rprecs <- c('rprec_SJ', 'rprec_caH', 'rprec_caL')
rprec_all <- join_dccs(rprecs)

# Add custom names for aggregation variables

rprec_all <- rprec_all %>% 
  mutate(month_name = case_when(
      month == "sep...AUG" ~ "Hydrol", 
      TRUE ~ as.character(.$month)))
```

## Plot precipitation 
```{r rwi_clim_prec}
p_prec <- rprec_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rprec_all$id, labels = rprec_all$month_name) + 
  theme(legend.position = c(.1, .8)) +
  annotate("text", label = "Prec", x = 1, y= -0.4) 
p_prec

pdf(paste0(di, '/out/climate_rwi/prec.pdf'), width=9, height = 5)
p_prec
dev.off()
```

# SPEI

* Correlación entre Marzo del mes anterior y hasta Sep del mes en curso
* Correlación con datos de spei6 hydrol, winter, summer, spring, autumn,

```{r}
set.seed(3333)
rspei_SJ <- dcc(chrono = cro_sj, climate = spei_N, 
         method = "correlation",  boot = "std", var_names = 'spei6',
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rspei_caL <- dcc(chrono = cro_caL, climate = spei_N, 
         method = "correlation",  boot = "std", var_names = 'spei6',
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rspei_caH <- dcc(chrono = cro_caH, climate = spei_N, 
         method = "correlation",  boot = "std", var_names = 'spei6',
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

# Joins dccs 
rspeis <- c('rspei_SJ', 'rspei_caH', 'rspei_caL')
rspei_all <- join_dccs(rspeis)

# Add custom names for aggregation variables of  and / or SPEI
customNames <- function(x){ 
  out <- x %>% 
    mutate(month_name = case_when(
      month == "sep...AUG" ~ "Hydrol",
      month == "dec...FEB" ~ "Winter",
      month == "MAR...MAY" ~ "Spring",
      month == "JUN...AUG" ~ "Summer",
      month == "SEP...NOV" ~ "Autumn",
      TRUE ~ as.character(.$month)))

  return(out)
}

rspei_all <- customNames(rspei_all) 
```

## Plot SPEI
```{r rwi_clim_spei}
p_spei <- rspei_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rspei_all$id, labels = rspei_all$month_name) + 
  theme(legend.position = c(.1, .8)) +
  annotate("text", label = "SPEI 6", x = 1, y= -0.4) 
p_spei

pdf(paste0(di, '/out/climate_rwi/spei.pdf'), width=9, height = 5)
p_spei
dev.off()
```

# Tmean

* Correlación entre Marzo del mes anterior y hasta Sep del mes en curso
* Correlación con tmean winter, summer, spring, autumn, tmean hydrol

```{r}
set.seed(3333)
rtmean_SJ <- dcc(chrono = cro_sj, climate = tmean_N, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmean_caH <- dcc(chrono = cro_caH, climate = tmean_S, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmean_caL <- dcc(chrono = cro_caL, climate = tmean_S, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

# Joins dccs 
rtmeans <- c('rtmean_SJ', 'rtmean_caH', 'rtmean_caL')
rtmean_all <- join_dccs(rtmeans)

# Add custom names for aggregation variables of Temperatures
customNames <- function(x){ 
  out <- x %>% 
    mutate(month_name = case_when(
      month == "sep...AUG" ~ "Hydrol",
      month == "dec...FEB" ~ "Winter",
      month == "MAR...MAY" ~ "Spring",
      month == "JUN...AUG" ~ "Summer",
      month == "SEP...NOV" ~ "Autumn",
      TRUE ~ as.character(.$month)))

  return(out)
  }

rtmean_all <- customNames(rtmean_all) 
```

## Plot Tmean 
```{r rwi_clim_tmean}
p_tmean <- rtmean_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rtmean_all$id, labels = rtmean_all$month_name) + 
  theme(legend.position ="top") + 
  annotate("text", label = "Tmean", x = 1, y= -0.4) + ylim(-.4, .4)
p_tmean

pdf(paste0(di, '/out/climate_rwi/tmean.pdf'), width=9, height = 5)
p_tmean
dev.off()
```

# Tmin

* Correlación entre Marzo del mes anterior y hasta Sep del mes en curso
* Correlación con tmin winter, summer, spring, autumn, tmean hydrol

```{r}
set.seed(3333)
rtmin_SJ <- dcc(chrono = cro_sj, climate = tmin_N, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmin_caH <- dcc(chrono = cro_caH, climate = tmin_S, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmin_caL <- dcc(chrono = cro_caL, climate = tmin_S, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

# Join dccs 
rtmins <- c('rtmin_SJ', 'rtmin_caH', 'rtmin_caL')
rtmin_all <- join_dccs(rtmins)

# Rename variable 
rtmin_all <- customNames(rtmin_all) 
```

## Plot Tmin 
```{r rwi_cli_tmin}
p_tmin <- rtmin_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rtmin_all$id, labels = rtmin_all$month_name) + 
  theme(legend.position = "none") +
  annotate("text", label = "Tmin", x = 1, y= -0.4) + ylim(-.4, .4)
p_tmin

pdf(paste0(di, '/out/climate_rwi/tmin.pdf'), width=9, height = 5)
p_tmin
dev.off()
```

# Tmax

* Correlación entre Marzo del mes anterior y hasta Sep del mes en curso
* Correlación con tmax winter, summer, spring, autumn, hydrol, may-jul, june-july

```{r}
set.seed(3333)
rtmax_SJ <- dcc(chrono = cro_sj, climate = tmax_N, 
         method = "correlation",  boot = "std", 
        .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11) + .mean(5:7) + .mean(6:7))

rtmax_caH <- dcc(chrono = cro_caH, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11) + .mean(5:7) + .mean(6:7))

rtmax_caL <- dcc(chrono = cro_caL, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-3:9) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11) + .mean(5:7) + .mean(6:7))

# Join dccs 
rtmaxs <- c('rtmax_SJ', 'rtmax_caH', 'rtmax_caL')
rtmax_all <- join_dccs(rtmaxs)

# Rename variable 
rtmax_all <- customNames(rtmax_all) 
```

## Plot Tmax 
```{r rwi_cli_tmax}
p_tmax <- rtmax_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rtmin_all$id, labels = rtmin_all$month_name) +
  theme(legend.position = "none") +
  annotate("text", label = "Tmax", x = 1, y= -0.4) + ylim(-.4, .4)
p_tmax

pdf(paste0(di, '/out/climate_rwi/tmax.pdf'), width=9, height = 5)
p_tmax
dev.off()
```

```{r rwi_cli_temperatures}
p_temp <- p_tmean + p_tmax + p_tmin + plot_layout(ncol=1)
pdf(paste0(di, '/out/climate_rwi/temp.pdf'), width=9, height = 12)
p_temp
dev.off()
```

```{r rwi_cli_prec_spei}
p_prec_spei <- p_prec + p_spei + plot_layout(ncol=1)
pdf(paste0(di, '/out/climate_rwi/prec_spei.pdf'), width=9, height = 8)
p_prec_spei
dev.off()
```

```{r rwi_cli_all}
p <- p_tmean + p_tmax + p_tmin + p_prec + plot_layout(ncol=2)
pdf(paste0(di, '/out/climate_rwi/all.pdf'), width=13, height = 8)
p
dev.off()
```

# Correlations RUNNING

Para explorar la variabilidad temporal (estabilidad) en la relación entre clima y crecimiento, utilizamos una correlación con window size de 30 años

## Prec 
```{r moving_precSJ}
set.seed(3333)
mprec_SJ <- dcc(chrono = cro_sj, climate = prec_N, selection =.sum(-9:8) + .range(3,12), 
         dynamic = "moving", win_size = 10, win_offset = 5, sb = FALSE, method = "cor", boot = 'std')
plot(mprec_SJ)
```

```{r moving_preccaH}
set.seed(3333)
mprec_caH <- dcc(chrono = cro_caH, climate = prec_S, selection =.sum(-9:8) + .range(3,12), 
         dynamic = "moving", win_size = 10, win_offset = 5, sb = FALSE, method = "cor", boot = 'std')
plot(mprec_caH)
```

```{r moving_preccaL}
set.seed(3333)
mprec_caL <- dcc(chrono = cro_caL, climate = prec_S, selection =.sum(-9:8) + .range(3,12), 
         dynamic = "moving", win_size = 10, win_offset = 5, sb = FALSE, method = "cor", boot = 'std')
plot(mprec_caL)
```
